---
title: ""
pagetitle: "TMU Pitcher Report"
output:
  html_document:
params:
  app_dir: "."
  rds_file: "tmu_session_data.rds"
  session_key: ""
  pitcher: ""
  team_name: "THO_MOR"
  session_label: "Live ABs"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = 'center')
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)

# ============================ ABOUT =================================
# TMU live-AB / scrimmage version of YallsPitcherReport.Rmd. Same front
# page and page-2 pitch log, rebuilt on the Offseason App's data:
#
#   - reads tmu_session_data.rds (the $liveab part), or the frame the app
#     hands over from "Download Pitcher Report" — overrides applied
#   - pitch types are the app's PitchGroup (hand tag / TrackMan fill /
#     manual override), never re-derived here
#   - PA results, AVG, K, BB etc. come from add_pa_outcomes() /
#     summarise_pa() in tmu_session_data.R, which this file SOURCES, so the
#     report and the app's Game Stats tab cannot disagree
#   - "vs. Opponent" is gone (it's all TMU); the Notes-based Base Situation
#     table is replaced by a vs LHH / vs RHH split
#   - one card per PITCHER PER SESSION, oldest first
#
# Standalone knit: put this file in the Offseason App folder, leave
# session_key blank, knit -> it lists the sessions; paste a SessionKey
# (or "all") into the params and knit again. pitcher narrows it to one or
# more arms: "Hanna, Brogan; Smith, Jake".
#
# Fixes carried over from the Y'alls version while converting it:
#   - Strike% was missing plain "FoulBall"; uses the app's STRIKE_CALLS
#   - Chase% counted untracked pitches as out-of-zone; now located only
#   - fastball threshold group said "Fastball" but tags are "Four-Seam",
#     so fastballs were being graded on the fallback thresholds
#   - Whiff% / Chase% printed 0% with no swings; now a dash
#   - Race-to-2K only caught 0-2 in two pitches; now 0-2 or 1-2 within
#     the first three pitches, or the PA over in three
# ====================================================================

# ============================ KNOBS =================================
SHOW_PITCH_LOG <- TRUE

# Outings longer than this print as two side-by-side half-tables.
LOG_SPLIT_MIN  <- 62

# Height in points the log TABLE should fill (title + legend sit outside).
LOG_TARGET_PT  <- 600

LOG_COLS_FULL    <- c("#","Inn","Batter","S","Cnt","Pitch","Res","Velo","Spin",
                      "iVB","HB","Ext","RelH","VAA","pX","pZ","EV","LA","Dist")
LOG_COLS_COMPACT <- c("#","Inn","Batter","S","Cnt","Pitch","Res","Velo","Spin",
                      "iVB","HB","EV","LA")

# Front-page density scale by pitch-type count. PAGE_FIT multiplies all of
# it: this card carries one more row than the Y'alls card (the game line
# under the header), so it starts a notch smaller. Nudge it if a print
# preview spills to a second page (down) or leaves a gap at the bottom (up).
PAGE_FIT <- 0.95
pt_scale_for <- function(n) {
  base <- if (is.na(n) || n <= 4) 1.00 else if (n == 5) 0.94 else if (n == 6) 0.88 else 0.82
  base * PAGE_FIT
}

# Zone plots in CATCHER'S VIEW, same as the app card (positive PlateLocSide
# — a RHH's inside — draws on the LEFT).
CATCHER_VIEW <- TRUE

# Whiff / Z-Whiff / Chase highlight thresholds by pitch family.
fastball_group <- c("Four-Seam", "Two-Seam", "Sinker")
breaking_group <- c("Cutter", "Slider", "Curveball", "Sweeper")
offspeed_group <- c("Changeup", "Splitter", "Knuckleball")

get_threshold <- function(pitch_type, metric) {
  if (pitch_type %in% fastball_group) {
    switch(metric, `Whiff %` = 17, `Chase %` = 18, `Z-Whiff %` = 15, 65)
  } else if (pitch_type %in% breaking_group) {
    switch(metric, `Whiff %` = 27, `Chase %` = 26, `Z-Whiff %` = 17, 65)
  } else if (pitch_type %in% offspeed_group) {
    switch(metric, `Whiff %` = 29, `Chase %` = 24, `Z-Whiff %` = 18, 65)
  } else {
    switch(metric, `Whiff %` = 25, `Chase %` = 25, `Z-Whiff %` = 17, 65)
  }
}
# ====================================================================

# ---- the app's data layer -------------------------------------------
app_dir <- normalizePath(params$app_dir, mustWork = FALSE)
helpers <- file.path(app_dir, "tmu_session_data.R")
if (!file.exists(helpers)) {
  stop("tmu_session_data.R not found in app_dir (", app_dir, "). Keep this ",
       ".Rmd in the Offseason App folder, or point app_dir at it.")
}
source(helpers, local = knitr::knit_global())

flip_name <- function(x) {
  vapply(strsplit(as.character(x), ",\\s*"), function(p) {
    if (length(p) >= 2 && !is.na(p[2])) paste(p[2], p[1]) else p[1]
  }, character(1))
}
disp_side <- function(ps) if (isTRUE(CATCHER_VIEW)) -ps else ps

split_keys <- function(x) {
  if (is.null(x) || !length(x)) return(character(0))
  v <- trimws(unlist(strsplit(as.character(x), ",")))
  v[nzchar(v)]
}
# "Last, First" contains a comma, so names are separated with ; or |
split_names <- function(x) {
  if (is.null(x) || !length(x) || !nzchar(x)) return(character(0))
  v <- trimws(unlist(strsplit(as.character(x), "[;|]")))
  v[nzchar(v)]
}

home_plate_segments <- data.frame(
  x    = c(0.71, 0.71,  0,   -0.71, -0.71),
  y    = c(0,    0.3,   0.5,  0.3,   0),
  xend = c(0.71, 0,    -0.71,-0.71,  0.71),
  yend = c(0.3,  0.5,   0.3,  0,     0)
)

# ---- read -----------------------------------------------------------
rds_path <- params$rds_file
if (!grepl("^([/~]|[A-Za-z]:)", rds_path)) rds_path <- file.path(app_dir, rds_path)
if (!file.exists(rds_path)) {
  stop("rds_file not found: ", rds_path,
       " — run build_tmu_session_data.R in the Offseason App folder first.")
}
obj <- readRDS(rds_path)

# A bare data frame is what the app hands over: already the selected sessions,
# so there is no picker. The build RDS is a list with $liveab.
from_app <- is.data.frame(obj)
full_data <- if (from_app) obj else obj$liveab
if (is.null(full_data) || !nrow(full_data)) {
  stop("No live-AB rows in ", basename(rds_path),
       if (!from_app && !is.null(obj$status$liveab))
         paste0(" (", obj$status$liveab, ")") else "")
}
full_data <- as.data.frame(full_data, stringsAsFactors = FALSE)

# Overrides: the app applies them before handing the frame over. A
# standalone knit applies tag_overrides.csv from the app folder itself.
if (!"IsOverride" %in% names(full_data)) {
  full_data <- apply_tag_overrides(
    full_data, read_tag_overrides(file.path(app_dir, TAG_OVERRIDE_FILE)))
}

# PA results on the WHOLE session frame, before any pitcher filter.
if (!"PAKey" %in% names(full_data)) full_data <- add_pa_outcomes(full_data)
if (!"Tracked" %in% names(full_data)) full_data$Tracked <- !is.na(full_data$RelSpeed)

# Team filter — skipped with a message if it would empty the report.
if (nzchar(params$team_name) &&
    any(full_data$PitcherTeam == params$team_name, na.rm = TRUE)) {
  full_data <- full_data[!is.na(full_data$PitcherTeam) &
                           full_data$PitcherTeam == params$team_name, , drop = FALSE]
} else if (nzchar(params$team_name)) {
  message("team_name '", params$team_name, "' matched no rows — keeping all.")
}

full_data$PlotPitchType <- as.character(full_data$PitchGroup)

sessions_index <- build_session_index(full_data) %>%
  transmute(SessionKey, Date, Pitches, Arms)

# ---- which sessions -------------------------------------------------
keys <- split_keys(params$session_key)
if (from_app || any(tolower(keys) == "all")) {
  session_chosen <- TRUE
} else if (length(keys)) {
  full_data <- full_data[full_data$SessionKey %in% keys, , drop = FALSE]
  session_chosen <- nrow(full_data) > 0
} else {
  session_chosen <- FALSE
}

# ---- which pitchers, one card per pitcher per session ---------------
want <- split_names(params$pitcher)
if (length(want)) full_data <- full_data[full_data$Pitcher %in% want, , drop = FALSE]

units <- if (session_chosen && nrow(full_data)) {
  full_data %>%
    group_by(SessionKey, Pitcher) %>%
    summarise(first_date  = {v <- Date[!is.na(Date)]
                             if (length(v)) min(v) else NA_character_},
              first_pitch = suppressWarnings(min(PitchNo, na.rm = TRUE)),
              .groups = "drop") %>%
    arrange(first_date, SessionKey, first_pitch, Pitcher) %>%
    as.data.frame(stringsAsFactors = FALSE)
} else data.frame()
```

```{r session_picker, echo=FALSE, results='asis'}
if (!session_chosen) {
  cat("## Pick a session\n\n")
  cat("Set **`session_key`** in the params header to one of the SessionKeys",
      "below (comma-separate several, or `all`), then re-knit.\n\n")
  print(knitr::kable(sessions_index, format = "html", table.attr = "class='table'"))
  knitr::knit_exit()
}
if (!nrow(units)) {
  cat("## Nothing to report\n\n")
  cat("The session / pitcher params matched no pitches.\n\n")
  knitr::knit_exit()
}
```
<style>
* {
  -webkit-print-color-adjust: exact !important;
  print-color-adjust: exact !important;
}

body {
  font-family: Arial, sans-serif;
  font-size: 11pt;
}
h1 {
  font-size: 20pt;
  margin-bottom: 5px;
  text-align: center;
  margin-top: 0;
}
h2 {
  font-size: 14pt;
  margin-top: 5px;
  margin-bottom: 5px;
  border-bottom: 2px solid #333;
}
.header-info {
  text-align: center;
  font-size: 12pt;
  margin-bottom: 5px;
}
table {
  font-size: 10pt;
}
.pitcher-section {
  page-break-after: always;
}
.pitcher-section:last-child {
  page-break-after: avoid;
}
.game-line table { margin: 2px auto 4px auto !important; }
.game-line th { background: #1a1a2e !important; color: #fff !important; }
.game-line td { font-weight: bold; }
.two-column-tables {
  display: flex;
  gap: 20px;
  margin-top: 5px;
}
.two-column-tables > div {
  flex: 1;
  width: 50%;
}
.two-column-plots {
  display: flex;
  gap: 20px;
  margin-top: 5px;
  align-items: flex-start;
}
.two-column-plots > div {
  flex: 1;
  width: 50%;
}
.three-column-plots {
  display: flex;
  gap: 20px;
  margin-top: 5px;
  align-items: flex-start;
}
.three-column-plots > div {
  flex: 1;
  width: 33.33%;
}
.empty-note { text-align: center; color: #666; font-style: italic; margin: 6px 0; }
/* per-pitcher density scale; overwritten inline as style="--pt: 0.80" */
.pitcher-section { --pt: 1; }

/* ---------------- page 2: pitch-by-pitch log ---------------- */
/* pandoc merges the raw pitcher-section div with its own H1 section div, so
   the log can end up NESTED inside .pitcher-section instead of a sibling.
   break-before on the log itself fires either way — do not rely on the
   parent's page-break-after. */
.pitch-log-section { --lf: 7pt; --lp: 2pt;
                     page-break-before: always; break-before: page; }
.log-split { page-break-inside: avoid; break-inside: avoid; }
.pitch-log-section h1 { font-size: 16pt; margin: 0 0 2px 0; text-align: center; }
.pitch-log-section .log-sub { text-align: center; font-size: 10pt;
                              color: #333; margin-bottom: 6px; }
.pitch-log-section table { font-size: var(--lf) !important; width: 100% !important;
                           margin: 0 auto !important; }
.pitch-log-section table td { padding: var(--lp) 2px !important; line-height: 1.12 !important;
                              white-space: nowrap !important; }
.pitch-log-section table th { padding: calc(var(--lp) + 1pt) 2px !important;
                              line-height: 1.12 !important;
                              white-space: nowrap !important;
                              font-size: var(--lf) !important; }
.log-split { display: flex; gap: 12px; align-items: flex-start; }
.log-split > div { flex: 1; width: 50%; }
.log-legend { font-size: 8pt; color: #444; margin-top: 6px; text-align: center; }

@media print {
  @page { size: letter portrait; margin: 0.3in; }

  html, body { margin: 0; padding: 0; font-size: 11pt; }

  .pitcher-section { page-break-after: always; page-break-inside: auto; }
  .pitcher-section:last-child { page-break-after: avoid; }

  h1 { font-size: 20pt; margin: 0 0 6px 0; page-break-after: avoid; }
  h2 { font-size: 12.5pt; margin: 4px 0 2px 0; page-break-after: avoid;
       border-bottom: 2px solid #333; }
  .header-info { font-size: 11pt; margin-bottom: 4px; }

  table { font-size: 8.6pt !important; margin: 0 auto 4px auto; width: 100% !important; }
  table td { padding: 3px 5px !important; line-height: 1.2 !important;
              white-space: nowrap !important; }
  table th { padding: 4px 4px !important; line-height: 1.15 !important;
             white-space: normal !important; }
  thead { display: table-header-group; }
  tr, img { page-break-inside: avoid; }
  .two-column-tables { page-break-inside: auto; }

  .two-column-tables, .two-column-plots, .three-column-plots {
    gap: 14px; margin-top: 5px;
  }

  .three-column-plots img { width: 100%; height: auto; max-height: 1.70in; }
  .two-column-plots  img { width: 100%; height: auto; max-height: 2.40in; }
  img { max-width: 100%; }

  .pitcher-section table    { font-size: calc(8.6pt * var(--pt)) !important; }
  .pitcher-section table td { padding: calc(3px * var(--pt)) calc(5px * var(--pt)) !important; }
  .pitcher-section table th { padding: calc(4px * var(--pt)) calc(4px * var(--pt)) !important; }
  .pitcher-section h1 { font-size: calc(20pt * var(--pt)); }
  .pitcher-section h2 { font-size: calc(12.5pt * var(--pt)); }
  .pitcher-section .header-info { font-size: calc(11pt * var(--pt)); }
  .pitcher-section .three-column-plots img { max-height: calc(1.70in * var(--pt)); }
  .pitcher-section .two-column-plots  img { max-height: calc(2.40in * var(--pt)); }
  .pitcher-section .two-column-tables,
  .pitcher-section .two-column-plots,
  .pitcher-section .three-column-plots { gap: calc(14px * var(--pt));
                                         margin-top: calc(5px * var(--pt)); }

  .pitch-log-section { page-break-before: always; break-before: page;
                       page-break-after: always; page-break-inside: auto; }
  .log-split { page-break-inside: avoid; break-inside: avoid; }
  .pitch-log-section:last-child { page-break-after: avoid; }
  .pitch-log-section h1 { font-size: 15pt; page-break-after: avoid; }
  .pitch-log-section table { font-size: var(--lf) !important; }
  .pitch-log-section table td { padding: var(--lp) 2px !important; }
  .pitch-log-section table th { padding: calc(var(--lp) + 1pt) 2px !important;
                                font-size: var(--lf) !important; }
  .log-split { gap: 10px; }
}
</style>
```{r results='asis'}
kbl_std <- function(x, font = 11, escape = TRUE) {
  x %>%
    kable(format = "html", align = "c", escape = escape) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = TRUE, font_size = font) %>%
    row_spec(0, bold = TRUE, background = "#f0f0f0")
}

emit_pair <- function(t1, k1, t2, k2) {
  cat('<div class="two-column-tables">\n<div>\n')
  if (!is.null(k1)) { cat(paste0('<h2>', t1, '</h2>\n')); print(k1) }
  cat('</div>\n<div>\n')
  if (!is.null(k2)) { cat(paste0('<h2>', t2, '</h2>\n')); print(k2) }
  cat('</div>\n</div>\n\n')
}

highlight_cell <- function(v, pitch_type, metric) {
  thr <- vapply(pitch_type, get_threshold, numeric(1), metric = metric)
  hot <- !is.na(v) & v >= thr
  cell_spec(ifelse(is.na(v), "\u2014", paste0(v, "%")), "html",
            bold = hot, color = ifelse(hot, "white", "black"),
            background = ifelse(hot, "#D30000", "transparent"))
}

for (u in seq_len(nrow(units))) {
  pitcher  <- units$Pitcher[u]
  sess_key <- units$SessionKey[u]

  data <- full_data %>%
    filter(Pitcher == pitcher, SessionKey == sess_key) %>%
    arrange(is.na(PitchNo), PitchNo)

  pitcher_name <- flip_name(pitcher)
  session_date <- pretty_session_date(units$first_date[u])
  # A game against another school reads "vs NKU"; a scrimmage keeps the label.
  opp <- if ("Opponent" %in% names(data)) data$Opponent[1] else NA_character_
  unit_label <- if (is.na(opp) || !nzchar(opp)) params$session_label else
    paste0("vs ", team_display(opp))

  hand <- unique(data$PitcherThrows); hand <- hand[!is.na(hand)]
  hand_lbl <- if (!length(hand)) "" else if (grepl("^L", hand[1])) "LHP" else "RHP"

  # Plots and rate stats use tracked pitches; results use every pitch.
  df_plot <- data %>% filter(Tracked, !is.na(RelSpeed))

  n_pt <- df_plot %>%
    filter(!is.na(PlotPitchType)) %>%
    distinct(PlotPitchType) %>% nrow()
  pt <- pt_scale_for(n_pt)

  pa   <- pa_table(data)
  line <- summarise_pa(pa)
  disc <- summarise_pitches(data)
  n_untr <- sum(!data$Tracked, na.rm = TRUE)

  cat(sprintf('<div class="pitcher-section" style="--pt: %.2f">\n\n', pt))
  cat(paste0("# ", pitcher_name, " | ", session_date, " | ",
             unit_label, "\n\n"))
  cat('<div class="header-info">',
      paste(c(hand_lbl,
              paste0(nrow(data), " Pitches",
                     if (n_untr > 0) paste0(" (", n_untr, " untracked)") else ""),
              paste0(line$PA, " BF"),
              paste0(n_pt, " Pitch Types")),
            collapse = " &nbsp;|&nbsp; "),
      '</div>\n', sep = "")

  # ================== GAME LINE ==================
  game_line <- tibble(
    BF = line$PA, AB = line$AB, H = line$H, `2B` = line$`2B`,
    `3B` = line$`3B`, HR = line$HR, BB = line$BB, HBP = line$HBP,
    K = line$K, AVG = fmt_avg(line$AVG), OBP = fmt_avg(line$OBP),
    SLG = fmt_avg(line$SLG), `K%` = fmt_pct(line$`K%`, 1),
    `BB%` = fmt_pct(line$`BB%`, 1))
  cat('<div class="game-line">\n')
  game_line %>%
    kable(format = "html", align = "c") %>%
    kable_styling(bootstrap_options = "condensed", full_width = TRUE, font_size = 11) %>%
    print()
  cat('</div>\n')
  cat('<hr>\n\n')

  if (nrow(df_plot) == 0) {
    cat('<div class="empty-note">No tracked pitches in this outing \u2014 ',
        'results above and the pitch log still count every charted pitch.</div>\n\n')
  } else {

  # ================== VELO DENSITY ==================
  mean_velo <- df_plot %>%
    group_by(PlotPitchType) %>%
    summarise(mean_speed = round(mean(RelSpeed, na.rm = TRUE), 1), .groups = "drop") %>%
    arrange(desc(mean_speed))

  df_plot$PlotPitchType   <- factor(df_plot$PlotPitchType, levels = mean_velo$PlotPitchType)
  mean_velo$PlotPitchType <- factor(mean_velo$PlotPitchType, levels = mean_velo$PlotPitchType)

  min_max_velo <- df_plot %>%
    group_by(PlotPitchType) %>%
    summarise(min_speed = round(min(RelSpeed, na.rm = TRUE), 1),
              max_speed = round(max(RelSpeed, na.rm = TRUE), 1), .groups = "drop")

  velo_density_plot <- ggplot(df_plot,
      aes(x = RelSpeed, color = PlotPitchType, fill = PlotPitchType)) +
    geom_density(alpha = 1, fill = NA, show.legend = FALSE, trim = TRUE) +
    geom_density(alpha = 0.4, color = NA, show.legend = FALSE, trim = TRUE) +
    facet_wrap(~PlotPitchType, ncol = if (n_pt >= 5) 2 else 1, dir = "v",
               scales = "free_y", strip.position = "left") +
    scale_color_manual(values = pitch_colors) +
    scale_fill_manual(values = pitch_colors) +
    geom_text(data = mean_velo,
              aes(x = mean_speed, y = 0, label = sprintf("%.1f", mean_speed)),
              vjust = -0.5, color = "black", size = if (n_pt >= 5) 2.6 else 3) +
    geom_vline(data = min_max_velo, aes(xintercept = min_speed),
               color = "black", linetype = "dashed") +
    geom_vline(data = min_max_velo, aes(xintercept = max_speed),
               color = "black", linetype = "dashed") +
    theme(
      axis.ticks.y = element_blank(),
      axis.line.y  = element_blank(),
      axis.text.y  = element_blank(),
      strip.text.y.left = element_text(angle = 0, size = if (n_pt >= 5) 10 else 14),
      panel.grid.major.x = element_line(color = "gray", linewidth = 0.1),
      panel.grid.minor.y = element_blank(),
      panel.background = element_rect(fill = "white"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
    ) +
    labs(title = "Velo Density", x = "Velocity", y = NULL) +
    theme(plot.title = element_text(size = 14, face = "bold"))

  # ================== USAGE LHH vs RHH ==================
  freq_data <- df_plot %>%
    filter(!is.na(BatterSide), !is.na(PlotPitchType)) %>%
    mutate(Side = case_when(
      trimws(BatterSide) %in% c("Left", "L", "left")   ~ "LHH",
      trimws(BatterSide) %in% c("Right", "R", "right") ~ "RHH",
      TRUE ~ NA_character_)) %>%
    filter(!is.na(Side)) %>%
    group_by(Side, PlotPitchType) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(Side) %>%
    mutate(freq = n / sum(n)) %>%
    ungroup() %>%
    mutate(freq_plot   = ifelse(Side == "LHH", -freq, freq),
           label       = paste0(round(freq * 100, 0), "%"),
           label_hjust = ifelse(Side == "LHH", 1.05, -0.05))

  if (nrow(freq_data) == 0) {
    usage_plot <- ggplot() + theme_void() +
      labs(title = "No Pitch Usage Data") +
      theme(plot.title = element_text(hjust = 0.5, face = "bold"))
  } else {
    usage_plot <- ggplot(freq_data,
                         aes(y = PlotPitchType, x = freq_plot, fill = PlotPitchType)) +
      geom_col(width = 0.6) +
      geom_vline(xintercept = 0, color = "black", linewidth = 1) +
      geom_text(aes(label = label, hjust = label_hjust),
                size = if (n_pt >= 5) 4.4 else 6, fontface = "bold", color = "black") +
      scale_x_continuous(labels = function(x) paste0(abs(x * 100), "%"),
                         limits = c(-0.9, 0.9)) +
      scale_fill_manual(values = pitch_colors, guide = "none") +
      labs(title = "Pitch Usage", subtitle = "LHH vs RHH", x = NULL, y = NULL) +
      theme_minimal(base_size = 13) +
      theme(plot.title = element_text(hjust = 0.5, face = "bold"),
            plot.subtitle = element_text(hjust = 0.5),
            panel.grid.major.y = element_blank(),
            panel.grid.minor = element_blank(),
            axis.text.y = element_blank(),
            legend.position = "none")
  }

  # ================== MOVEMENT ==================
  df_movement <- df_plot %>% filter(!is.na(HorzBreak), !is.na(InducedVertBreak))

  movement_plot <- ggplot(df_movement,
      aes(x = HorzBreak, y = InducedVertBreak, color = PlotPitchType)) +
    geom_point(size = 2.5, alpha = 0.7) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    geom_vline(xintercept = 0, linetype = "dashed") +
    scale_color_manual(values = pitch_colors) +
    scale_x_continuous(limits = c(-24, 24), breaks = seq(-24, 24, 6)) +
    scale_y_continuous(limits = c(-24, 24), breaks = seq(-24, 24, 6)) +
    labs(title = "Pitch Movement Plot",
         x = "Horizontal Break (inches)",
         y = "Induced Vertical Break (inches)",
         color = "Pitch Type") +
    coord_fixed() +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold", size = 14),
          legend.position = "bottom")

  cat('<div class="three-column-plots">\n')
  cat('<div>\n\n'); print(velo_density_plot); cat('\n\n</div>\n')
  cat('<div>\n\n'); print(usage_plot);        cat('\n\n</div>\n')
  cat('<div>\n\n'); print(movement_plot);     cat('\n\n</div>\n')
  cat('</div>\n\n')

  # ================== VELO BY INNING (or by batter) ==================
  # Live ABs are often one long "inning". With fewer than two innings
  # charted, the x axis becomes the batter faced instead, so the plot
  # still shows whether velo held up.
  n_inn <- dplyr::n_distinct(df_plot$Inning[!is.na(df_plot$Inning)])
  if (n_inn >= 2) {
    velo_x <- df_plot %>% mutate(X = Inning)
    x_lab  <- "Inning"; v_title <- "Avg Velocity by Inning"
  } else {
    bf_order <- unique(data$PAKey)
    velo_x <- df_plot %>% mutate(X = match(PAKey, bf_order))
    x_lab  <- "Batter Faced"; v_title <- "Avg Velocity by Batter Faced"
  }
  velo_inning_data <- velo_x %>%
    filter(!is.na(X)) %>%
    group_by(X, PlotPitchType) %>%
    summarise(avg_velo = round(mean(RelSpeed, na.rm = TRUE), 1), .groups = "drop")

  velo_inning_plot <- ggplot(velo_inning_data,
      aes(x = X, y = avg_velo, color = PlotPitchType, group = PlotPitchType)) +
    geom_line(linewidth = 1) +
    geom_point(size = 3) +
    geom_text(aes(label = avg_velo), vjust = -0.8, size = 3, fontface = "bold",
              show.legend = FALSE) +
    scale_color_manual(values = pitch_colors) +
    scale_x_continuous(breaks = sort(unique(velo_inning_data$X))) +
    labs(title = v_title, x = x_lab, y = "Avg Velocity (mph)", color = "Pitch Type") +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold", size = 14),
          legend.position = "bottom",
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "white"),
          panel.border = element_rect(color = "black", fill = NA, linewidth = 1))

  # ================== ZONE BY PITCH TYPE ==================
  zone_df <- df_plot %>%
    filter(!is.na(PlateLocSide), !is.na(PlateLocHeight)) %>%
    mutate(PX = disp_side(PlateLocSide))
  pitch_counts  <- zone_df %>% count(PlotPitchType, name = "n_pitches")
  heatmap_types <- pitch_counts %>% filter(n_pitches >= 100) %>% pull(PlotPitchType)
  point_types   <- pitch_counts %>% filter(n_pitches <  100) %>% pull(PlotPitchType)

  zone_plot <- ggplot() +
    geom_rect(data = data.frame(xmin = SZ_XMIN, xmax = SZ_XMAX,
                                ymin = SZ_YMIN, ymax = SZ_YMAX),
              aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
              inherit.aes = FALSE, fill = NA, color = "black", linewidth = 1.2) +
    geom_segment(data = home_plate_segments,
                 aes(x = x, y = y, xend = xend, yend = yend),
                 color = "black", linewidth = 1, inherit.aes = FALSE) +
    stat_density_2d(
      data = zone_df %>% filter(PlotPitchType %in% heatmap_types),
      aes(x = PX, y = PlateLocHeight, fill = after_stat(level)),
      geom = "polygon", alpha = 0.45, contour = TRUE, show.legend = FALSE) +
    scale_fill_gradient(low = "#ffe5e5", high = "#cc0000") +
    geom_point(
      data = zone_df %>% filter(PlotPitchType %in% point_types),
      aes(x = PX, y = PlateLocHeight, color = PlotPitchType),
      alpha = 0.8, size = 2) +
    facet_wrap(~PlotPitchType, ncol = if (n_pt <= 4) 2 else 3) +
    scale_color_manual(values = pitch_colors, guide = "none") +
    scale_y_continuous(limits = c(0, 5.5)) +
    scale_x_continuous(limits = c(-2.5, 2.5)) +
    labs(title = "Strike Zone Plot by Pitch Type",
         subtitle = if (isTRUE(CATCHER_VIEW)) "Catcher's view" else "Pitcher's view",
         x = NULL, y = NULL) +
    coord_fixed() +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold", size = 14),
          plot.subtitle = element_text(size = 9, color = "gray40"),
          legend.position = "none",
          strip.text = element_text(face = "bold", size = if (n_pt >= 5) 9 else 11),
          panel.grid.major = element_line(color = "gray90"),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "white"),
          panel.border = element_rect(color = "black", fill = NA, linewidth = 1))

  cat('<div class="two-column-plots">\n')
  cat('<div>\n\n'); print(velo_inning_plot); cat('\n\n</div>\n')
  cat('<div>\n\n'); print(zone_plot);        cat('\n\n</div>\n')
  cat('</div>\n\n')
  }  # end tracked-pitch plots

  # ================== PITCH SUMMARY ==================
  # Flags come from add_session_flags() — the same definitions the app uses.
  # AVG is by the pitch the PA ENDED on.
  flagged <- add_session_flags(df_plot) %>%
    mutate(IsZWhiff    = HasLoc & InZone & IsWhiff,
           IsZoneSwing = HasLoc & InZone & IsSwing,
           IsOZ        = HasLoc & !InZone,
           PA_AB  = !is.na(PAEnd) & PAEnd & PAResult %in% AB_RESULTS,
           PA_H   = !is.na(PAEnd) & PAEnd & PAResult %in% HIT_RESULTS,
           PA_K   = !is.na(PAEnd) & PAEnd & PAResult == "K",
           PA_BB  = !is.na(PAEnd) & PAEnd & PAResult == "BB")

  summary_table <- flagged %>%
    group_by(PlotPitchType) %>%
    summarise(
      `#`        = n(),
      `Avg Velo` = round(mean(RelSpeed, na.rm = TRUE), 1),
      `Max Velo` = round(max(RelSpeed, na.rm = TRUE), 1),
      `Spin`     = round(mean(SpinRate, na.rm = TRUE), 0),
      `iVB`      = round(mean(InducedVertBreak, na.rm = TRUE), 1),
      `HB`       = round(mean(HorzBreak, na.rm = TRUE), 1),
      # fmt_num() dashes a missing mean — Yakkertech has no extension, and
      # paste0() on NaN printed "NaNft".
      `RelH`     = paste0(fmt_num(mean(RelHeight, na.rm = TRUE), 1), "ft"),
      `Ext`      = paste0(fmt_num(mean(Extension, na.rm = TRUE), 1), "ft"),
      `VAA`      = paste0(fmt_num(mean(VertApprAngle, na.rm = TRUE), 1), "\u00B0"),
      `Zone %`   = ifelse(sum(HasLoc) == 0, "\u2014",
                          paste0(round(100 * sum(InZone) / sum(HasLoc), 1), "%")),
      `Strike %` = paste0(round(100 * mean(IsStrike, na.rm = TRUE), 1), "%"),
      `Whiff %`  = ifelse(sum(IsSwing) == 0, NA_real_,
                          round(100 * sum(IsWhiff) / sum(IsSwing), 1)),
      `Z-Whiff %`= ifelse(sum(IsZoneSwing) == 0, NA_real_,
                          round(100 * sum(IsZWhiff) / sum(IsZoneSwing), 1)),
      `Chase %`  = ifelse(sum(IsOZ) == 0, NA_real_,
                          round(100 * sum(IsChase) / sum(IsOZ), 1)),
      `AVG`      = fmt_avg(.rate(sum(PA_H), sum(PA_AB))),
      `K/BB`     = paste0(sum(PA_K), "/", sum(PA_BB)),
      .groups = "drop") %>%
    rename(`Pitch Type` = PlotPitchType) %>%
    mutate(`Pitch Type` = as.character(`Pitch Type`),
           across(c(RelH, Ext, VAA), ~ sub("^\u2014.*$", "\u2014", .x)))

  if (nrow(summary_table)) {
    summary_table <- summary_table %>%
      mutate(`Whiff %`   = highlight_cell(`Whiff %`,   `Pitch Type`, "Whiff %"),
             `Z-Whiff %` = highlight_cell(`Z-Whiff %`, `Pitch Type`, "Z-Whiff %"),
             `Chase %`   = highlight_cell(`Chase %`,   `Pitch Type`, "Chase %"))

    cat('\n\n## Pitch Summary\n\n')
    summary_table %>%
      kable(format = "html", align = "c", escape = FALSE) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                    full_width = TRUE, font_size = 11) %>%
      row_spec(0, bold = TRUE, background = "#f0f0f0") %>%
      print()
    cat('\n\n')
  }

  # ================== BATTED BALL SUMMARY ==================
  bip <- flagged %>% filter(IsBIP)
  bb_kbl <- if (nrow(bip) == 0) NULL else {
    bip %>%
      mutate(BBT = batted_ball_type(Angle)) %>%
      group_by(PlotPitchType) %>%
      summarise(
        `BIP`    = n(),
        `H`      = sum(PAResult %in% HIT_RESULTS & PAEnd, na.rm = TRUE),
        `Avg EV` = fmt_num(mean(ExitSpeed, na.rm = TRUE), 1),
        `GB %`   = fmt_pct(.rate(sum(BBT == "GB", na.rm = TRUE), sum(!is.na(BBT))), 0),
        `LD %`   = fmt_pct(.rate(sum(BBT == "LD", na.rm = TRUE), sum(!is.na(BBT))), 0),
        `FB %`   = fmt_pct(.rate(sum(BBT == "FB", na.rm = TRUE), sum(!is.na(BBT))), 0),
        `Hard Hit %` = fmt_pct(.rate(sum(ExitSpeed >= HARD_HIT_MPH, na.rm = TRUE),
                                     sum(!is.na(ExitSpeed))), 0),
        .groups = "drop") %>%
      rename(`Pitch Type` = PlotPitchType) %>%
      kbl_std()
  }
  if (is.null(bb_kbl)) {
    bb_kbl <- tibble(Note = "No balls in play") %>% kbl_std()
  }

  # ================== PLATOON SPLITS ==================
  # Replaces the Y'alls Base Situation table: TMU live ABs don't carry the
  # hand-typed base state in Notes, but every PA has a batter side.
  side_of <- function(x) case_when(
    trimws(x) %in% c("Left", "L", "left")   ~ "vs LHH",
    trimws(x) %in% c("Right", "R", "right") ~ "vs RHH",
    TRUE ~ NA_character_)
  plat_rows <- lapply(c("vs LHH", "vs RHH"), function(sd) {
    p_sd <- if (is.null(pa)) NULL else pa[side_of(pa$BatterSide) %in% sd, , drop = FALSE]
    d_sd <- data[side_of(data$BatterSide) %in% sd, , drop = FALSE]
    if (!nrow(d_sd)) return(NULL)
    s1 <- summarise_pa(p_sd); s2 <- summarise_pitches(d_sd)
    tibble(Split = sd, P = nrow(d_sd), PA = s1$PA, AVG = fmt_avg(s1$AVG),
           OBP = fmt_avg(s1$OBP), K = s1$K, BB = s1$BB,
           `Strike %` = fmt_pct(s2$`Strike%`), `Whiff %` = fmt_pct(s2$`Whiff%`),
           `Chase %` = fmt_pct(s2$`Chase%`))
  })
  plat_tbl <- bind_rows(plat_rows)
  plat_kbl <- if (nrow(plat_tbl)) kbl_std(plat_tbl, font = 9) else NULL

  # ================== USAGE BY COUNT ==================
  usage_kbl <- NULL
  usage_cnt <- df_plot %>%
    filter(!is.na(Balls), !is.na(Strikes)) %>%
    mutate(Count = case_when(
      Balls == 0 & Strikes == 0 ~ "0-0",
      Strikes == 2              ~ "2K",
      Strikes > Balls           ~ "Ahead",
      Balls > Strikes           ~ "Behind",
      TRUE                      ~ "Even"),
      Count = factor(Count, levels = c("0-0", "Ahead", "Even", "Behind", "2K")))
  if (nrow(usage_cnt)) {
    usage_kbl <- usage_cnt %>%
      mutate(PlotPitchType = as.character(PlotPitchType)) %>%
      group_by(Count, PlotPitchType) %>%
      summarise(n = n(), .groups = "drop_last") %>%
      mutate(pct = paste0(round(100 * n / sum(n), 0), "%")) %>%
      ungroup() %>% select(-n) %>%
      tidyr::pivot_wider(names_from = PlotPitchType, values_from = pct,
                         values_fill = "\u2014") %>%
      arrange(Count) %>%
      kbl_std(font = 9)
  }

  # ================== OVERALL METRICS ==================
  # PA-level: FPS%, Race-to-2K, Putaway% are over HIS completed PAs.
  pa_pitches <- data %>% filter(!is.na(PAKey)) %>% group_by(PAKey) %>%
    summarise(Reached2K = any(!is.na(Strikes) & Strikes == 2),
              EarlyTwo  = any(!is.na(Strikes) & Strikes == 2 &
                                !is.na(Balls) & Balls <= 1 & PitchInPA <= 4),
              .groups = "drop")
  pa_x <- if (is.null(pa)) NULL else left_join(pa, pa_pitches, by = "PAKey")

  r2k_pct <- if (is.null(pa_x) || !nrow(pa_x)) NA_real_ else
    mean(pa_x$EarlyTwo | (pa_x$Pitches <= 3 & !pa_x$PAResult %in% c("BB", "HBP")))
  putaway_pct <- if (is.null(pa_x) || !any(pa_x$Reached2K)) NA_real_ else
    mean(pa_x$PAResult[pa_x$Reached2K] %in% OUT_RESULTS)

  # Whole percents and the 9pt table font: nine columns share half a page,
  # and "100.0%" at 11pt ran BB% off the right edge.
  overall_metrics <- tibble(
    `FPS%`     = fmt_pct(disc$`FPS%`),
    `R2K%`     = fmt_pct(r2k_pct),
    `Putaway%` = fmt_pct(putaway_pct),
    `Whiff%`   = fmt_pct(disc$`Whiff%`),
    `Chase%`   = fmt_pct(disc$`Chase%`),
    `CSW%`     = fmt_pct(disc$`CSW%`),
    `Soft%`    = fmt_pct(disc$`Soft%`),
    `K%`       = fmt_pct(line$`K%`),
    `BB%`      = fmt_pct(line$`BB%`))
  om_kbl <- kbl_std(overall_metrics, font = 9)

  emit_pair('Batted Ball Summary', bb_kbl,    'Platoon Splits',  plat_kbl)
  emit_pair('Usage by Count',      usage_kbl, 'Overall Metrics', om_kbl)

  cat('</div>\n\n')

  # ==================================================================
  # PAGE 2 — PITCH-BY-PITCH LOG ("back of the sheet")
  # ------------------------------------------------------------------
  # Built from `data`, NOT df_plot, so untracked pitches still appear in
  # the sequence and the count matches the real outing. The PA-ending
  # pitch shows the PA RESULT from add_pa_outcomes() — so a K inferred from
  # the count prints as K, not CS.
  # ==================================================================
  if (SHOW_PITCH_LOG && nrow(data) > 0) {

    log_pitch_colors <- pitch_colors
    log_pitch_colors[c("Cutter","Changeup","Sweeper","Knuckleball")] <-
      c("#B7950B","#1E8449","#0E7490","#4D7A2E")

    fmt1 <- function(x) ifelse(is.na(.num(x)), "\u2014", sprintf("%.1f", .num(x)))
    fmt0 <- function(x) ifelse(is.na(.num(x)), "\u2014", sprintf("%.0f", .num(x)))
    short_name <- function(x) {
      out <- vapply(strsplit(as.character(x), ",\\s*"), function(p) {
        if (length(p) >= 2 && !is.na(p[2])) paste0(p[1], ", ", substr(p[2], 1, 1))
        else if (length(p) >= 1) p[1] else NA_character_
      }, character(1))
      ifelse(is.na(out), "\u2014", out)
    }

    log_raw <- data %>%
      mutate(
        .inn = .num(Inning),
        .pc  = ifelse(is.na(PitchCall), "", as.character(PitchCall)),
        .ppt = as.character(PlotPitchType),
        .newpa = PAKey != dplyr::lag(PAKey, default = ""),
        .pitch_res = case_when(
          .pc == "HitByPitch"                         ~ "HBP",
          .pc %in% BALL_CALLS                         ~ "B",
          .pc == "StrikeCalled"                       ~ "CS",
          .pc == "StrikeSwinging"                     ~ "SW",
          grepl("^Foul", .pc)                         ~ "F",
          .pc == "InPlay"                             ~ "IP",
          TRUE                                        ~ "\u2014"),
        .res = case_when(
          !is.na(PAEnd) & PAEnd & PAResult %in% c("SF", "SH") ~ "SAC",
          !is.na(PAEnd) & PAEnd & PAResult == "BIP"           ~ "IP",
          !is.na(PAEnd) & PAEnd & PAResult != "Incomplete"    ~ PAResult,
          TRUE                                                ~ .pitch_res),
        .res_col = case_when(
          .res %in% c("B","BB","HBP")          ~ "#C0392B",
          .res %in% c("CS","K","SW")           ~ "#1E8449",
          .res == "F"                          ~ "#B9770E",
          .res %in% c("1B","2B","3B","HR")     ~ "#7D3C98",
          TRUE                                 ~ "#21618C"),
        .res_bold = .res %in% c("K","BB","HBP","SW","HR","3B","2B","1B"),
        .bat = short_name(Batter),
        .batshort = substr(sub(",.*$", "", ifelse(is.na(Batter), "\u2014",
                                                  as.character(Batter))), 1, 10),
        .side = ifelse(is.na(BatterSide), "", substr(as.character(BatterSide), 1, 1))
      )

    log_disp <- log_raw %>%
      transmute(
        `#`    = dplyr::row_number(),
        Inn    = ifelse(is.na(.inn), "\u2014", as.character(.inn)),
        Batter = ifelse(.newpa, .bat, ""),
        S      = ifelse(.newpa, .side, ""),
        Cnt    = ifelse(is.na(.num(Balls)) | is.na(.num(Strikes)), "\u2014",
                        paste0(as.integer(.num(Balls)), "-", as.integer(.num(Strikes)))),
        Pitch  = cell_spec(
                   ifelse(!is.na(.ppt) & .ppt %in% names(pitch_abbr),
                          unname(pitch_abbr[.ppt]),
                          ifelse(is.na(.ppt), "\u2014", substr(.ppt, 1, 2))),
                   "html", bold = TRUE,
                   color = ifelse(!is.na(.ppt) & .ppt %in% names(log_pitch_colors),
                                  unname(log_pitch_colors[.ppt]), "#333333")),
        Res    = cell_spec(.res, "html", bold = .res_bold, color = .res_col),
        Velo   = fmt1(RelSpeed),
        Spin   = fmt0(SpinRate),
        iVB    = fmt1(InducedVertBreak),
        HB     = fmt1(HorzBreak),
        Ext    = fmt1(Extension),
        RelH   = fmt1(RelHeight),
        VAA    = fmt1(VertApprAngle),
        pX     = fmt1(PlateLocSide),
        pZ     = fmt1(PlateLocHeight),
        EV     = ifelse(.pc == "InPlay", fmt1(ExitSpeed), "\u2014"),
        LA     = ifelse(.pc == "InPlay", fmt0(Angle),     "\u2014"),
        Dist   = ifelse(.pc == "InPlay", fmt0(Distance),  "\u2014"),
        .inn      = .inn,
        .newpa    = .newpa,
        .bat      = .bat,
        .batshort = .batshort,
        .side     = .side
      )

    render_log_tbl <- function(d, cols) {
      iv    <- ifelse(is.na(d$.inn), -1, d$.inn)
      grp   <- cumsum(c(TRUE, diff(iv) != 0))
      shade <- which(grp %% 2 == 0)
      brdr  <- which(d$.newpa) - 1
      brdr  <- brdr[brdr >= 1]
      both  <- intersect(shade, brdr)
      s_o   <- setdiff(shade, brdr)
      b_o   <- setdiff(brdr, shade)
      bcss  <- "border-bottom: 1px solid #aab2ba;"

      k <- d[, cols, drop = FALSE] %>%
        kable(format = "html", align = "c", escape = FALSE) %>%
        kable_styling(bootstrap_options = c("condensed"), full_width = TRUE) %>%
        row_spec(0, bold = TRUE, background = "#e6e9ec")
      if (length(s_o))  k <- row_spec(k, s_o,  background = "#eef2f6")
      if (length(b_o))  k <- row_spec(k, b_o,  extra_css = bcss)
      if (length(both)) k <- row_spec(k, both, background = "#eef2f6", extra_css = bcss)
      k
    }

    n_log     <- nrow(log_disp)
    split_log <- n_log > LOG_SPLIT_MIN
    cols_use  <- if (split_log) LOG_COLS_COMPACT else LOG_COLS_FULL
    rows_side <- if (split_log) ceiling(n_log / 2) else n_log
    lf <- if (!split_log) {
            if (rows_side <= 30) 9.6 else if (rows_side <= 45) 8.8 else 7.8
          } else if (rows_side <= 55) 7.2
            else if (rows_side <= 70) 6.8
            else 6.2
    text_pt <- rows_side * lf * 1.12
    lp <- round(min(3.8, max(0.4, (LOG_TARGET_PT - text_pt) / (rows_side + 1) / 2)), 2)

    if (split_log) {
      log_disp$Batter <- ifelse(log_disp$.newpa, log_disp$.batshort, "")
    }

    n_strikes <- sum(log_raw$.pc %in% STRIKE_CALLS, na.rm = TRUE)

    cat(sprintf('<div class="pitch-log-section" style="--lf: %.1fpt; --lp: %.2fpt">\n\n',
                lf, lp))
    cat(paste0('<h1>', pitcher_name, ' &mdash; Pitch by Pitch</h1>\n'))
    cat(paste0('<div class="log-sub">', session_date, ' &nbsp;|&nbsp; ',
               unit_label, ' &nbsp;|&nbsp; ', n_log, ' P &nbsp;|&nbsp; ',
               line$PA, ' BF &nbsp;|&nbsp; ',
               line$H, ' H &nbsp;|&nbsp; ', line$K, ' K &nbsp;|&nbsp; ',
               line$BB, ' BB &nbsp;|&nbsp; ',
               round(100 * n_strikes / n_log, 0), '% strikes</div>\n'))

    if (split_log) {
      half <- ceiling(n_log / 2)
      d1 <- log_disp[1:half, , drop = FALSE]
      d2 <- log_disp[(half + 1):n_log, , drop = FALSE]
      if (identical(d2$Batter[1], "")) {
        d2$Batter[1] <- d2$.batshort[1]
        d2$S[1]      <- d2$.side[1]
      }
      cat('<div class="log-split">\n<div>\n\n')
      print(render_log_tbl(d1, cols_use))
      cat('\n\n</div>\n<div>\n\n')
      print(render_log_tbl(d2, cols_use))
      cat('\n\n</div>\n</div>\n\n')
    } else {
      print(render_log_tbl(log_disp, cols_use))
    }

    cat(paste0('<div class="log-legend">',
               'B = ball &middot; CS = called strike &middot; SW = swinging strike &middot; ',
               'F = foul &middot; K / BB / HBP = PA ended &middot; ',
               '1B / 2B / 3B / HR / Out / FC / E / SAC / IP = ball in play ',
               '(IP = no result charted) &middot; Cnt = count before the pitch',
               '</div>\n'))
    cat('</div>\n\n')
  }
}
```
